DGE's and Enrichment

### 2022-01-31 reports/rmarkdown/manuscript/Figure2.Rmd
update meta data points

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

par(mar=c(4,.1,4,.2)) #b,l,t,r
h <- hclust( as.dist( (1- cor(t(top_picrust_data)))/2 ), method = "ward.D2")
plot( rev(as.dendrogram(h)) ,xlim=c(max(h$height),-1), horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
title(xlab="Height", line=2)
cutoff <- 4
abline(v=cutoff,xpd=F,col="red",lty=2)
gene_module <- cutree(h, h = cutoff)

@@

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
mtext("                        Differentially expressed KEGG Ontology terms and pathways", 
      side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
title(xlab="Study Groups", line=1)

end <- table(pal[factor(group_annotation)[ order(group_annotation[sample_use])]] )
end <- end[order(factor(names(end), levels = pal))]
end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/ncol(picrust_data[sample_use])-.015) )
start <- c(0, map_dbl(end[1:4], ~.x+0.015))

axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
# text labels
t <- map2(start, end, ~(.x+.y)/2)
map2(t,levels(factor(group_annotation)), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

points( seq(0,1,length.out = length(sample_use)),
        rep(par("usr")[4],length(sample_use))+.005, pch=15, cex=.7,xpd=T,
        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]])
points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.008, pch=15, xpd=T,cex=.5,
        col= c("tomato","orange","#d7d7d7")[as.factor(metadata$BV_Diagnosis_v3)][order( group_annotation[sample_use])] )
points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.024, pch=15, xpd=T,cex=.5,
        col= c("#d7d7d7","tomato")[as.factor(metadata$HIVstatus)][order(group_annotation[sample_use])])

@@

temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
df <- data.frame(temp,temp)
quietly(plot_sankey)( df, pal = taxa_pal[as.numeric(levels(temp))], use_w2 = F, plot_labels = T,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)$output

>>>>

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(scales)
  library(igraph)
  library(RColorBrewer)
  #remotes::install_github("czarnewski/niceRplots",force=T)
  library(niceRplots)
})  })  })

@@

# Dendogram
par(mar=c(4,.1,4,.2)) #b,l,t,r
h <- hclust( as.dist( (1- cor(t(top_picrust_data)))/2 ), method = "ward.D2")
plot( rev(as.dendrogram(h)) ,xlim=c(max(h$height),-1), horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
title(xlab="Height", line=2)
cutoff <- 4
abline(v=cutoff,xpd=F,col="red",lty=2)
gene_module <- cutree(h, h = cutoff)

@@

# Heatmap
image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
mtext("                        Differentially expressed KEGG Ontology terms and pathways", 
      side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
title(xlab="Study Groups", line=1)

# Study Group brackets
end <- table(pal[factor(group_annotation)[ order(group_annotation[sample_use])]] )
end <- end[order(factor(names(end), levels = pal))]
end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/ncol(picrust_data[sample_use])-.015) )
start <- c(0, map_dbl(end[1:4], ~.x+0.017))
end <- c(end[1:length(end)-1], par("usr")[2]-.005)

axis(1, at = c(start, end), label = F, pos =1.021 , xpd=T, col="white",col.ticks="black")
map2(start, end, ~lines(x=c(.y, .x), y= c(1.021, 1.021), xpd=T, cex=.8))
# text labels
t <- map2(start, end, ~(.x+.y)/2)
map2(t,levels(factor(group_annotation)), ~text(x=.x, y=1.037, .y,xpd=T, cex=.8))

# Meta bar info
par(mar=c(3,.1,2,.1)) #bottom, left, top, right  
map(c(.0021, .006), ~points( seq(0,1,length.out = length(sample_use)),
        rep(par("usr")[4],length(sample_use))+.x, pch=15, cex=.3,xpd=F,
        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]]) )
points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.008, pch=15, xpd=F,cex=.3,
        col= c("tomato","orange","#d7d7d7")[as.factor(metadata$BV_Diagnosis_v3)][order( group_annotation[sample_use])] )
points( seq(0,1,length.out = ncol(top_picrust_data)), 
        rep(par("usr")[3],ncol(top_picrust_data))-.024, pch=15, xpd=F,cex=.3,
        col= c("#d7d7d7","tomato")[as.factor(metadata$HIVstatus)][order(group_annotation[sample_use])])

@@

temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
df <- data.frame(temp,temp)
par(mar=c(4,0,4,.1)) #bottom, left, top, right  
quietly(plot_sankey)( df, pal = taxa_pal[as.numeric(levels(temp))], use_w2 = F, plot_labels = T,gapv = .01,gap2v = 0 ,xaxs="i",yaxs="i",plot_weights = F)$output

### 2022-01-31 reports/rmarkdown/manuscript/Figure4-5.Rmd
update meta data points

  ## Heatmap
  par(mar=c(4,.1,2.5,.1))
  plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
  title(xlab="Height", line=2)
  abline(v=cuttoff,xpd=F,col="red",lty=2)
  points( rep(-.1,length(gene_module)),
          seq(length(gene_module),1,length.out = length(gene_module)),
          col=taxa_pal[factor(gene_module[h$order])],
          pch=15,cex=.5,xpd=F)

@@

  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("                                                          Differentially expressed genes and KEGG/GO pathways", side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

  end <- table(pal[factor(names(samples))[ order(df[1])]])
  end <- end[order(factor(names(end), levels = pal))]
  end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
  start <- c(0, map_dbl(end[1:4], ~.x+0.015))

  axis(1, at = c(start, end), label = F, pos =1.027 , xpd=T, col="white",col.ticks="black")
  map2(start, end, ~lines(x=c(.y, .x), y= c(1.027, 1.027), xpd=T, cex=.8))
  # text labels
  t <- map2(start, end, ~(.x+.y)/2)
  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.043, .y,xpd=T, cex=.8))

  ## Meta info bars
  points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.005,
          col=pal[factor(names(samples))[ order(df[1])]],
          pch=15,xpd=T,cex=.7)
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
          col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45) 
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
          col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
          pch=15,xpd=T,cex=.45)

@@

  temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
  plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
               use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",
               yaxs="i",plot_weights = F)

  text(par("usr")[1],par("usr")[3]-.008,labels = "    BV",cex=.6, xpd=T)
  text(par("usr")[1],par("usr")[3]-.028,labels = "    HIV",cex=.6, xpd=T)

@@

# test figure
# dev.copy2pdf(file=paste0("./test.pdf"),
#     width = 6.6,
#     height = 6, paper = "a4"
#     )
```

@@

sheets <- excel_sheets(paste0(out, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)
```

>>>>

  # Dendogram
  par(mar=c(4,.1,2.5,.1)) #bottom, left, top, right  
  plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T)
  title(xlab="Height", line=2)
  abline(v=cuttoff,xpd=F,col="red",lty=2)
  points( rep(-.1,length(gene_module)),
          seq(length(gene_module),1,length.out = length(gene_module)),
          col=taxa_pal[factor(gene_module[h$order])],
          pch=15,cex=.5,xpd=F)

@@

  # Heatmap
  image( t(top_TRX_counts[h$order,pull(df, "ID")][nrow(top_TRX_counts):1,]),
         col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
         breaks = seq(-5,5,length.out = 100),axes=F)
  mtext("                                                          Differentially expressed genes and KEGG/GO pathways", side=3, las = 1, cex = .7, line =1.5,xpd=T, font = 2)
  title(xlab="Study Groups", line=1)

  # Study Group brackets
  end <- table(pal[factor(names(samples))[ order(df[1])]])
  end <- end[order(factor(names(end), levels = pal))]
  end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
  start <- c(0, map_dbl(end[1:4], ~.x+0.017))
  end <- c(end[1:length(end)-1], par("usr")[2]-.005)

  axis(1, at = c(start, end), label = F, pos =1.029 , xpd=T, col="white",col.ticks="black")
  map2(start, end, ~lines(x=c(.y, .x), y= c(1.029, 1.029), xpd=T, cex=.8))
  # text labels
  t <- map2(start, end, ~(.x+.y)/2)
  par(mar=c(3,.2,2,.2)) #bottom, left, top, right  
  map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.045, .y,xpd=T, cex=.8))

  ## Meta info bars
  par(mar=c(3,.1,2,.1)) #bottom, left, top, right  
  map(c(.0025, .007), ~points( seq(0,1,length.out = length(samples)),rep(par("usr")[4],length(samples))+.x,
          col=pal[factor(names(samples))[ order(df[1])]],
          pch=15,xpd=F,cex=.3) )
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.008,
          col=c("tomato","orange","#d7d7d7")[ factor(metadata[samples,"BV_Diagnosis_v3"][order(df[1])]) ],
          pch=15,xpd=F,cex=.3) 
  points( seq(0,1,length.out = ncol(top_TRX_counts)),rep(par("usr")[3],ncol(top_TRX_counts))-.022,
          col= c("#d7d7d7","tomato")[ factor(metadata[samples,"HIVstatus"][order(df[1])]) ],
          pch=15,xpd=F,cex=.3)

@@

  temp <- factor( rev(gene_module), levels = unique( gene_module[h$order] ))
  par(mar=c(4,0,2.5,.1)) #bottom, left, top, right  
  plot_sankey( data.frame(temp,temp), pal = taxa_pal[as.numeric(levels(temp))], 
               use_w2 = F , plot_labels = T ,gapv = .01,gap2v = 0 ,xaxs="i",
               yaxs="i",plot_weights = F)

  text(par("usr")[1]+0.008,par("usr")[3]-.008,labels = "     BV",cex=.6, xpd=T)
  text(par("usr")[1]+0.008,par("usr")[3]-.028,labels = "      HIV",cex=.6, xpd=T)

@@

# test figure
dev.copy2pdf(file=paste0("./test.pdf"),
    width = 6.6,
    height = 6, paper = "a4"
    )
```

@@

sheets <- excel_sheets(paste0(out, DEG_files[2])) %>% set_names(.) %>% .[-c(1:2)]
arrange_plots.fun(enrich_T, "Tissue_gr", "Tissue", sheets, "TF_PPI_enrichment", enrichR_files[2], Layout = l)

dev.copy2pdf(file=paste0("./test.pdf"),
    width = 6.6,
    height = 6, paper = "a4"
    )
```
